Stage-Splitting a Modal Language

### 2014-06-17 paper/semantics.tex
made a bunch of small fixes

The grammar and type system for our language, \lang, are given in
\ref{fig:lamStagedSyntax} and \ref{fig:coreStatics}.
We formulated our typing judgments in the style of \cite{davies96}, where the whole judgment is annotated with a stage.  
Variables in the context are annotated with a stage and must be used at the same stage at which they were declared.
All of the non-stage constructs have the expected type-level behavior.  
This is made manifest as rules which are entirely abstract over stage.

Note that $\next$ and $\prev$ are the introduction and elimination forms for $\fut$ types.
Specifically, $\next$, given an argument with type $A$ at stage \bbtwo, forms a $\fut A$ at stage \bbone.  
That is, it forms the promise of a future $A$ out of a construction for an $A$ at the next timestep.
This promise can only be ``cashed in" be the $\prev$ construct.  
Since $\prev$ operates at stage \bbtwo, this ensures no violation of causality.

@@

Fast exponentiation is a method for calculating $b^p$ in $\log p$ time:
\[
	\mathit{fe}(b,p) = \left \{ \begin{array}{ll} 
		1 &  e = 0 \\ 
		\mathit{fe}(b,p/2)^2 & p \text{ even} \\ 
		b \cdot \mathit{fe}(b,p-1) & p \text{ odd} \end{array}
	\right .
\]
It is often observed that the decomposition of the power depends only on the power at prior recursive levels.
For this reason, we can stage an implementation of fast exponentiation so that the power is available at the first stage and the base is available at the second stage.
This implies that the result is then available only at the second stage.

@@

\begin{lstlisting} 
let exp (b : $int, e : int) : $int =
	if e == 0 then
		next{1}
	else if (e mod 2) == 0 then
		next{let x = prev{exp(b,e/2)} in x*x}
	else
		next{prev{b} * prev{exp (b,e-1)}}		
\end{lstlisting}

@@

% FIXME but do we give a cost semantics?
Given this definition of fast exponent, the question the remains how it can be run.  For this, we need a semantics for \lang.

>>>>

The grammar and type system for our language, \lang, are given in
\ref{fig:grammar} and \ref{fig:statics}.
We formulated our typing judgments in the style of \cite{davies96}, where the whole judgment is annotated with a stage.  
Variables in the context are annotated with a stage and must be used at the same stage at which they were declared.
All of the non-stage constructs have the expected type-level behavior.  
This is made manifest as rules which are entirely abstract over stage.

In addition to determining the stage, $\next$ and $\prev$ are the introduction and elimination forms for $\fut$ types.
Specifically, $\next$, given an argument with type $A$ at stage \bbtwo, forms a $\fut A$ at stage \bbone.  
That is, it forms the promise of a future $A$ out of a construction for an $A$ at the next timestep.
This promise can only be redeemed be the $\prev$ construct.  
Since $\prev$ operates at stage \bbtwo, this ensures no violation of causality.

@@

Fast exponentiation is a method for calculating $b^p$ in $\log p$ time:
\[
	\mathit{fexp}(b,p) = \left \{ \begin{array}{ll} 
		1 &  p = 0 \\ 
		\mathit{fexp}(b,p/2)^2 & p \text{ even} \\ 
		b \cdot \mathit{fexp}(b,p-1) & p \text{ odd} \end{array}
	\right .
\]
It is often observed that the decomposition of the power depends only on the power at prior recursive levels.
For this reason, we can stage an implementation of fast exponentiation so that the power is available at the first stage and the base is available at the second stage.
This implies that the result is then available only at the second stage.

@@

\begin{lstlisting} 
let fexp (b : $int, e : int) : $int =
	if e == 0 then
		next{1}
	else if (e mod 2) == 0 then
		next{let x = prev{fexp(b,e/2)} in x*x}
	else
		next{prev{b} * prev{fexp(b,e-1)}}		
\end{lstlisting}

@@

% FIXME but do we give a cost semantics? no
Given this definition of fast exponent, the question the remains how it can be run.  For this, we need a semantics for \lang.
